### 2020-09-17 manuscript/ERL_global_C_review.Rmd
minor updates

On an ecosystem level, forest C cycling is central to the flow of energy and materials, and as the basis of food webs...
(support of biodiversity)
(human relience as source of food, fuel, fiber)

@@

This review synthesizes data from the ForC database [https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018]. ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. 
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. 
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. 
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018). 
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018]. 
Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v.##; Cook-Patton et al. in press), and data from several publications (Taylor et al. 2017, Lutz et al. 2018, Johnson_2018_csss; see GitHub repo for complete list) ([GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)). 
We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: XX).

@@

Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 2-5, and statistics for each biome type are also summarized at: [GitHub URL]. 
Of the # flux and # stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests. 
For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. 
That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig. 4). **update this: https://github.com/forc-db/ERL-review/issues/16 **

@@

(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 7-10. In general, ForC contained roughly comparable information on C fluxes in regrowth vs mature forests, with the notable exception of tropical forests, for which there were no fluxes that had been measured in $\ge$ 7 different geographic areas. C stocks were better documented for regrowth stands, with a total of # records from # distinct geographic areas. Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Fig. 6-11; detailed below). Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs. 5-6).

>>>>

On an ecosystem level, forest C cycling is central to the flow of energy and materials, and as the basis of food webs...
(support of biodiversity)
(human reliance as source of food, fuel, fiber)

@@

This review synthesizes data from the ForC database [https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018]. ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. 
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. 
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. 
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible (83% of records in v.2.0; Anderson-Teixeira et al 2018). 
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018]. 
Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v1.0; Cook-Patton et al. in press; ZenodoDOI), and data from several publications (Taylor et al. 2017, Lutz et al. 2018, Johnson_2018_csss; see GitHub repo for complete list) ([GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)). 
We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: TBD).

@@

Average C cycles for tropical broadleaf, temperate broadleaf, temperate conifer, and boreal forests $\ge$ 100 years old and with no known major disturbance or significant anthropogenic management are presented in Figures 1-4, and statistics for each biome type are also summarized at: [GitHub URL]. 
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, # fluxes and # stocks in temperate broadleaf forests, #fluxes and # stocks in temperate conifer forests, and fluxes and # stocks in boreal forests. 
For variables with records from $\ge$ 7 distinct geographic areas, these ensemble C budgets were generally consistent. 
That is, component fluxes and stocks summed to within 1 std of more inclusive fluxes in all but one instance (in temperate conifer forests, $aboveground woody biomass + foliage biomass > aboveground biomass + 1std$; Fig. 4). **update this: https://github.com/forc-db/ERL-review/issues/16 **

@@

(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 7-10. In general, ForC contained roughly comparable information on C fluxes in regrowth vs mature forests, with the notable exception of tropical forests, for which there were no fluxes that had been measured in $\ge$ 7 different geographic areas. C stocks were better documented for regrowth stands, with a total of # records from # distinct geographic areas. Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Fig. 5-10; detailed below). Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs. 5-6).
